Excel fun

### 2022-07-11 Excel fun.Rmd
Image explanation of long vs. wide data.
Headings and explanations.

@@

```{r}
test2 <- test |> 
  mutate(
    category = ifelse(is.na(value), entry, NA)
  ) |> 
  fill(category) |> 
  fill(year) |> 
  filter(!is.na(value))
test2
```

@@

```{r}
read_excel("Data/Book1.xlsx") |> 
  filter(!is.na(Year))
```

>>>>

## Long vs. wide data
![](Images/wideLong1-1.png)

Fra [Statology](https://www.statology.org/long-vs-wide-data/).

@@

```{r}
test2 <- test |> 
  mutate(
    category = ifelse(is.na(value), entry, NA) # transferring headings to category
  ) |> 
  fill(category) |> # filling category column
  fill(year) |> # filling year column
  filter(!is.na(value)) # removing headings
test2
```

### Removing rows containing 'sum' and 'total'
```{r}
test2_no_totals <- test2 |> 
  filter(
    !str_detect(entry, "^Sum") & !str_detect(entry, "^Total")
  )

test2_no_totals
test2_no_totals |>
  group_by(category) |> 
  summarise(
    sum = sum(value)
  )
```

@@

```{r}
wide_data <- read_excel("Data/Book1.xlsx")
wide_data
wide_data |> 
  filter(!is.na(Year))
```
